---
title: "analysis_iv_ii"
author: "Hunter York"
date: "10/31/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(ggplot2)
library(dplyr)
library(magrittr)
library(stringr)
library(readxl)
library(ggrepel)
library(foreign)
library(ipumsr)
library(data.table)
library(factoextra)



options
theme_set(theme_bw())

# load the data
setwd("/Users/hyork/Documents/projects/occupation/code")

weighted.var <- function(x, w, na.rm = FALSE) {
  if (na.rm) {
    w <- w[i <- !is.na(x)]
    x <- x[i]
  }
  sum.w <- sum(w)
  sum.w2 <- sum(w^2)
  mean.w <- sum(x * w) / sum(w)
  (sum.w / (sum.w^2 - sum.w2)) * sum(w * (x - mean.w)^2, na.rm =
                                       na.rm)
}



# acs <- readstata13::read.dta13("../inputs/usa_00012.dta")
# acs <- acs[acs$empstat == "employed" &
#                              (acs$school == "no, not in school"| as.numeric(as.character(acs$age)) >=34) &
#                              as.numeric(as.character(acs$age)) %in% 25:64 &
#                             acs$wkswork2 %like% "40|48|50" &
#                              acs$incwage != 999998 &
#                              acs$incwage > 0,]
# acs <- data.table(acs)
# #keep vars
# acs <- acs[,.(race, hispan, year,statefip, countyicp,metro,perwt, sex, age, school,educ, empstat,occ, ind,ind1950, classwkr, wkswork2, incwage, occsoc)]
# 
# # subset


#saveRDS(acs, "../inputs/usa_00012.rds")
acs <- readRDS( "../inputs/usa_00012.rds")

setnames(acs, c("occsoc", "ind1950"), c("origoccsoc", "origind1950"))
# derive a 5-year age var
acs[, age_cat := paste0(floor(as.numeric(as.character(age))/5)*5,
                        " to ", 
                        floor(as.numeric(as.character(age))/5)*5 + 4)]

# make occ var char
acs[, occ := as.character(occ)]

# 
acs[nchar(occ) == 1, occ := paste0("00", occ)]
acs[nchar(occ) == 2, occ := paste0("0", occ)]

# make first dig var
acs[, first_dig := substr(occ, 1,1)]

# make map for first dig categories
map <- data.table(first_dig = as.character(0:9), 
                  occ_categ = c("Professional", "Clerical", "Service",
                                "Agricultural, etc.", "Skilled", "Skilled",
                                "Semiskilled", "Semiskilled", "Unskilled",
                                "Unskilled"))
acs <- merge(acs, map, by = "first_dig")

# merge census 1940 file on ind codes
ind_codes <- read_xlsx("../ref/IND1940.xlsx") %>% data.table()
ind_codes[, IND := as.numeric(IND)]
acs <- merge(acs, 
             ind_codes[!is.na(IND) & IND != 996], 
             by.x = "ind", by.y = "IND", all.x = T)

# crosswalk to ppp
ppp <- read_excel("../ref/CPI_U_RS.xlsx") %>% data.table
ppp[1, year := 1940]
ppp[1, cpi := 33]
ppp[, year := as.character(year)]
ppp[, xwalk_fac := cpi/369.8]
acs <- merge(acs, ppp, by = "year")
acs[, incwage := incwage/xwalk_fac]
# create log inc wage
acs[, log_incwage := log(incwage + 1)]

# crosswalk to occsoc
occ_soc <- fread("../ref/occ_occsoc_crosswalk_2000_onward_without_code_descriptions.csv")
occ_soc[, seq := 1:nrow(occ_soc)]
occ_soc <- melt(occ_soc, id.vars = c("Occupation title", "V1", "seq"))
occ_soc[, pref := substr(variable, 1,4)]
occ_soc[, soc := ifelse(variable %like% "SOC"|variable == "2018 Onward ACS/PRCS", 1,0)]
occ_soc <- occ_soc[pref %in% c(2005, 2013, 2018)]
occ_soc <- dcast(occ_soc, `Occupation title` + pref+seq~ soc, value.var = "value")
setnames(occ_soc, c("soc_occ_title", "year", "seq", "occ", "OCCSOC"))
occ_soc <- occ_soc[!is.na(occ), .(soc_occ_title, year, occ, OCCSOC)]
occ_soc[year == 2005, year := 2009]

# copy less granular 2009 categories to 2013 and 2018
temp <- occ_soc[year == 2009 & !occ %in% unique(occ_soc[year == 2013, occ])]
temp[, year := 2013]
occ_soc <- rbind(occ_soc, temp)
temp <- occ_soc[year == 2009 & !occ %in% unique(occ_soc[year == 2018, occ])]
temp[, year := 2018]
occ_soc <- rbind(occ_soc, temp)
temp <- occ_soc[year == 2013 & !occ %in% unique(occ_soc[year == 2009, occ])]
temp[, year := 2009]
occ_soc <- rbind(occ_soc, temp)
temp <- occ_soc[year == 2013 & !occ %in% unique(occ_soc[year == 2018, occ])]
temp[, year := 2018]
occ_soc <- rbind(occ_soc, temp)

temp <- occ_soc[year == 2018 & !occ %in% unique(occ_soc[year == 2013, occ])]
temp[, year := 2013]
occ_soc <- rbind(occ_soc, temp)
temp <- occ_soc[year == 2018 & !occ %in% unique(occ_soc[year == 2009, occ])]
temp[, year := 2009]
occ_soc <- rbind(occ_soc, temp)

# merge on acs
acs[, occ := as.numeric(occ)]
occ_soc[, occ:= as.numeric(occ)]
acs <- merge(acs, occ_soc, by = c("occ", "year"), all.x = T)

acs[, OCCSOC := origoccsoc]

# now create skills dataset
skills_2009 <- read.delim('../ref/db_14_0 2009.7/Skills.txt') %>% data.table()
skills_2013 <-  read.delim('../ref/db_18_0_2013.7/Skills.txt') %>% data.table()
skills_2018 <- read_excel("../ref/db_22_2_excel 2018.2/Skills.xlsx") %>% data.table()
skills_2009[, year := 2009]
skills_2013[, year := 2013]
skills_2018[, year := 2018]
setnames(skills_2018, names(skills_2018), gsub(" ", ".", names(skills_2018), fixed = T))
skills_2018[, `O.NET.SOC.Code` := `O*NET-SOC.Code`]
skills <- rbindlist(list(skills_2009, skills_2013, skills_2018), fill = T)
skills <- skills[, .(O.NET.SOC.Code, Element.Name, Scale.ID, Data.Value, Standard.Error, year)]

# reformate onet codes to merge
skills[, OCCSOC := gsub("-", "", substr(O.NET.SOC.Code,1,7))]
skills <- skills[Scale.ID == "LV",.(Data.Value = mean(Data.Value),
                                    Standard.Error = mean(as.numeric(Standard.Error))), by = .(Element.Name,Scale.ID, OCCSOC)]

occ_soc_x <- acs[OCCSOC %like% "X", .(year, OCCSOC)] %>% unique() # these are wildcard census vars
setnames(occ_soc_x, "OCCSOC", "tempOCCSOC")

#merge on wild cards one digit at a time
for (i in 4:1){
  acs[, newvar := str_sub(OCCSOC, 1, -1-i)]
  occ_soc_x[, newvar := str_sub(tempOCCSOC, 1, -1-i)]
  acs <- merge(acs, occ_soc_x[tempOCCSOC %like% paste0("[0-9]", paste0(rep("X", i), collapse = ""), "$")] ,
               by = c("newvar", "year"), all.x = T)
  acs[!is.na(tempOCCSOC), OCCSOC := tempOCCSOC]
  acs[, newvar := NULL]
  acs[, tempOCCSOC := NULL]
  occ_soc_x[, newvar := NULL]
}

# some things aren't being merged. Fix later
#acs[!OCCSOC %in% unique(skills$OCCSOC), unique(OCCSOC)]
# for all things that end in a 0, make an average of children
acs[!OCCSOC %in% unique(skills$OCCSOC), unique(OCCSOC)] -> fixes
skills[, Standard.Error := as.numeric(Standard.Error)]
temp <- skills[substr(OCCSOC,1,5) %in% substr(fixes[fixes %like% "0$"],1,5)]
temp[, OCCSOC := paste0(substr(OCCSOC,1,5), "0")]
temp <- temp[,.(Data.Value = mean(Data.Value),
                Standard.Error = mean(Standard.Error)), by = .(Element.Name, Scale.ID, OCCSOC)]
skills <- rbind(temp, skills, fill = T)
# do the same for ending in 99
temp <- skills[substr(OCCSOC,1,4) %in% substr(fixes[fixes %like% "99$"],1,4)]
temp[, OCCSOC := paste0(substr(OCCSOC,1,4), "99")]
temp <- temp[,.(Data.Value = mean(Data.Value),
                Standard.Error = mean(Standard.Error)), by = .(Element.Name, Scale.ID, OCCSOC)]
skills <- rbind(temp, skills, fill = T)
# do the same for ending in XXXX
temp <- skills[substr(OCCSOC,1,2) %in% substr(fixes[fixes %like% "XXXX$"],1,2)& !substr(OCCSOC,1,2) %like% "X"]
temp[, OCCSOC := paste0(substr(OCCSOC,1,2), "XXXX")]
temp <- temp[,.(Data.Value = mean(Data.Value),
                Standard.Error = mean(Standard.Error)), by = .(Element.Name, Scale.ID, OCCSOC)]
skills <- rbind(temp, skills, fill = T)
# do the same for ending in XXX
temp <- skills[substr(OCCSOC,1,3) %in% substr(fixes[fixes %like% "XXX$"],1,3) & !substr(OCCSOC,1,3) %like% "X"]
temp[, OCCSOC := paste0(substr(OCCSOC,1,3), "XXX")]
temp <- temp[,.(Data.Value = mean(Data.Value),
                Standard.Error = mean(Standard.Error)), by = .(Element.Name, Scale.ID, OCCSOC)]
skills <- rbind(temp, skills, fill = T)
# do the same for ending in XX
temp <- skills[substr(OCCSOC,1,4) %in% substr(fixes[fixes %like% "XX$"],1,4)& !substr(OCCSOC,1,4) %like% "X"]
temp[, OCCSOC := paste0(substr(OCCSOC,1,4), "XX")]
temp <- temp[,.(Data.Value = mean(Data.Value),
                Standard.Error = mean(Standard.Error)), by = .(Element.Name, Scale.ID, OCCSOC)]
skills <- rbind(temp, skills, fill = T)
# do the same for ending in X
temp <- skills[substr(OCCSOC,1,5) %in% substr(fixes[fixes %like% "X$"],1,5)& !substr(OCCSOC,1,5) %like% "X"]
temp[, OCCSOC := paste0(substr(OCCSOC,1,5), "X")]
temp <- temp[,.(Data.Value = mean(Data.Value),
                Standard.Error = mean(Standard.Error)), by = .(Element.Name, Scale.ID, OCCSOC)]
skills <- rbind(temp, skills, fill = T)
# for all things that end in a 0, make an average of children
temp <- skills[substr(OCCSOC,1,5) %in% substr(fixes[fixes %like% "X$"],1,5)]
temp[, OCCSOC := paste0(substr(OCCSOC,1,5), "0")]
temp <- temp[,.(Data.Value = mean(Data.Value),
                Standard.Error = mean(Standard.Error)), by = .(Element.Name, Scale.ID, OCCSOC)]
skills <- rbind(temp, skills, fill = T)


#okay now one real final last try to fill in missing vars
acs[!OCCSOC %in% unique(skills$OCCSOC) & !is.na(OCCSOC), unique(OCCSOC)] -> fixes

for(c.occsoc in fixes){
  keeper <- "continue"
  for(i in 1:4){
    print(c.occsoc)
    print(i)
    if(keeper == "continue"){
      fixes_substr <- str_sub(c.occsoc,1,-1-i)
      candidates <- unique(skills$OCCSOC)
      matches <- candidates[str_sub(candidates, 1, -1-i) == fixes_substr]
      if(length(matches > 1)){
        temp <- skills[OCCSOC %in% matches]
        temp[, OCCSOC := c.occsoc]
        temp <- temp[,.(Data.Value = mean(Data.Value),
                        Standard.Error = mean(Standard.Error)), by = .(Element.Name, Scale.ID, OCCSOC)]
        skills <- rbind(temp, skills, fill = T)
        keeper <- "stop"
      }
    }
  }
}

# #create a map of all children
# parents <- data.table(parents = unique(skills$OCCSOC)[unique(skills$OCCSOC) %like% "X$|0$"])
# parents_out <- data.table()
# for(c.parent in parents$parents){
#     substr_parent <- gsub("0$|00$|000$|0000$|X$|XX$|XXX$|XXXX$", "", c.parent, perl = TRUE)
#     children <- data.table(parents = c.parent, children = unique(skills$OCCSOC)[unique(skills$OCCSOC) %like% paste0("^", substr_parent)])
#     temp <- merge(parents, children, allow.cartesian = T)
#     if(nrow(temp)>1){
#         parents_out <- rbind(parents_out, temp, fill = T)
#     }
# }
# parents_out[, sub :=  gsub("0$|00$|000$|0000$|X$|XX$|XXX$|XXXX$", "", parents, perl = TRUE)]
# parents_out[, level := 6-nchar(sub)]
# 
# # see how many children occsocs aren't in every year
# acs[,.(year, OCCSOC)] %>% unique -> all_children
# 
# # get least common set
# all_children[year == 2009, unique(OCCSOC)][all_children[year == 2009, unique(OCCSOC)] %in%
#                                                all_children[year == 2013, unique(OCCSOC)] & 
#                                                all_children[year == 2009, unique(OCCSOC)] %in% 
#                                                all_children[year == 2018, unique(OCCSOC)]] -> common_set
# 
# children_fixes <- parents_out[!children %in% common_set]
# children_fixes[,N := .N, by = .(children)]
# children_fixes[,maxN := max(N), by = .(children)]
# children_fixes[,maxlevel := max(level), by = .(children)]
# 
# children_fixes <- children_fixes[ N == 1 | level == maxlevel]
# 
# children_fixes <- children_fixes[!duplicated(children_fixes$children)]
#     
# acs <- merge(acs, children_fixes[,.(parents, children)], 
#              by.x = "OCCSOC", by.y = "children", all.x = T)
# acs[!is.na(parents), OCCSOC := parents]

#standardize by percent
skills <- skills[OCCSOC %in% unique(acs$OCCSOC)]

skills[,Data.Value := percent_rank(Data.Value), by = .(Element.Name, Scale.ID)]



```

# Examine how the import of skills has changed over time

## Do a quick PCA to see how skills vary with respect to each other

```{r}



# cast wide and do factor analysis
skills <- skills[OCCSOC %in% unique(acs$OCCSOC),
                 .(Data.Value = mean(Data.Value)), by = .(Element.Name, Scale.ID, OCCSOC)]


skills_wide <- dcast(skills[Scale.ID == "LV"], OCCSOC   ~Element.Name, value.var = "Data.Value")
res.pca <- prcomp(skills_wide[,2:34], scale = TRUE, center = T)

fviz_pca_var(res.pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE ,
             select.var = list(contrib = 15)    # Avoid text overlapping
)
fviz_pca_var(res.pca,
             axes = c(3,4),
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE,# Avoid text overlapping,
             select.var = list(contrib = 15)
)

```

## See how mean skills change across survey year using ACS data

```{r}

# make a skills dataset that has a few interesting variables
skills_sum <- skills[Scale.ID == "LV",
                     .(average_value_skills = mean(Data.Value, na.rm = T)), by = .(OCCSOC)]


acs <- merge(acs, skills_sum, by = c("OCCSOC"), all.x = T)

skills_overview <- acs[,.(mean_skills = weighted.mean(average_value_skills, w = perwt, na.rm = T)), by = .(year, age_cat, sex)]

skills_overview[, cohort := floor((as.numeric(year) - as.numeric(substr(age_cat,1,2)))/10)*10]

ggplot(skills_overview) + 
  geom_line(aes(x = year, y = mean_skills, color = age_cat, group = age_cat)) + 
  facet_wrap(~sex)

skills_overview[,.(mean_skills = mean(mean_skills)), by = .(cohort, year, sex)] %>% 
  ggplot(.) + 
  geom_line(aes(x = year, y = mean_skills, color = as.factor(cohort), group = as.factor(cohort))) + 
  facet_wrap(~sex) + 
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1))
```

## Recalculate skills by year to capture more interesting variables, like LV1 and LV2, Programming, etc

```{r}
# skills wide
skills_sum <- skills_wide
skills_sum[, pc1 := predict(res.pca, newdata = .SD)[,1], .SDcols = names(skills_sum)]
skills_sum[, pc2 := predict(res.pca, newdata = .SD)[,2], .SDcols = names(skills_sum)]
skills_sum[, pc3 := predict(res.pca, newdata = .SD)[,3], .SDcols = names(skills_sum)]
skills_sum[, pc4 := predict(res.pca, newdata = .SD)[,4], .SDcols = names(skills_sum)]

skills_sum[, programming := Programming]
skills_sum[, tech_skills := Programming + `Complex Problem Solving` +
             `Mathematics` + Programming + Science + `Systems Analysis` + 
             Troubleshooting]
skills_sum[, average_value_skills := rowMeans(.SD), .SDcols = rownames(res.pca$rotation)]


```


## See how mean skills change across survey year

```{r}
# average across years
# REMOVE THIS IF YOU WANT YEAR SPECIFIC SKILLS RATINGS
# skills_sum <- skills_sum[year == 2018,.(pc1 = mean(pc1), 
#                             pc2 = mean(pc2),
#                             programming = mean(programming)), by = "OCCSOC"]

acs <- merge(acs, skills_sum[,.(OCCSOC, pc1, pc2,pc3, pc4, programming, tech_skills)], by = c( "OCCSOC"), all.x = T)

# 

skills_overview2 <- acs[,.(pc1 = weighted.mean(pc1,w = perwt, na.rm = T),
                           pc2 = weighted.mean(pc2,w = perwt,  na.rm = T), 
                           pc3 = weighted.mean(pc3,w = perwt,  na.rm = T),
                           pc4 = weighted.mean(pc4,w = perwt,  na.rm = T),
                           programming = weighted.mean(programming,w = perwt,  na.rm = T),
                           tech_skills = weighted.mean(tech_skills, w = perwt, na.rm = T),
                           average_value_skills = weighted.mean(average_value_skills, w = perwt, na.rm = T)), by = .(year, age_cat, sex)]

skills_overview2_melt <- melt(skills_overview2, id.vars = c("year", "age_cat", "sex"))

skills_overview2_melt[, cohort := floor((as.numeric(year) - as.numeric(substr(age_cat,1,2)))/10)*10]


ggplot(skills_overview2_melt) + 
  geom_line(aes(x = year, y = value, color = age_cat, group = age_cat)) + 
  facet_grid(variable~sex, scales = "free") + 
  scale_color_viridis_d()+  
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1))

skills_overview2_melt[,.(value = mean(value)), by = .(cohort, sex, variable, year)] %>% 
  ggplot(.) + 
  geom_line(aes(x = year, y = value, color = as.factor(cohort), group = as.factor(cohort))) + 
  facet_grid(variable~sex, scales = "free") + 
  scale_color_viridis_d()+  
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1))
```

## See how changes in skill level are spread over occupational grouping

```{r}
skills_overview3 <- acs[,.(pc1 = weighted.mean(pc1,w = perwt, na.rm = T),
                           pc2 = weighted.mean(pc2,w = perwt,  na.rm = T), 
                           pc3 = weighted.mean(pc3,w = perwt,  na.rm = T),
                           pc4 = weighted.mean(pc4,w = perwt,  na.rm = T),
                           programming = weighted.mean(programming,w = perwt,  na.rm = T),
                           tech_skills = weighted.mean(tech_skills, w = perwt, na.rm = T), average_value_skills = weighted.mean(average_value_skills, w = perwt, na.rm = T)), by = .(year, age_cat, sex, occ_categ)]

skills_overview3_melt <- melt(skills_overview3, id.vars = c("year", "age_cat", "sex" ,"occ_categ"))

skills_overview3_melt[, cohort := floor((as.numeric(year) - as.numeric(substr(age_cat,1,2)))/10)*10]

skills_overview3_melt <- skills_overview3_melt[,.(value = mean(value)), by = .(cohort, variable, year, occ_categ, sex)]

ggplot(skills_overview3_melt[variable %like% "pc1|progr|tech|averag" &
                               occ_categ %in% unique(skills_overview3_melt$occ_categ)[1:3]]) + 
  geom_line(aes(x = year, y = value, color = as.factor(cohort), group = as.factor(cohort))) + 
  facet_grid( variable ~occ_categ+sex, scales = "free") + 
  scale_color_viridis_d()+  
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1))

ggplot(skills_overview3_melt[variable %like% "pc1|progr|tech|averag" &
                               occ_categ %in% unique(skills_overview3_melt$occ_categ)[4:7]]) + 
  geom_line(aes(x = year, y = value, color = as.factor(cohort), group = as.factor(cohort))) + 
  facet_grid( variable ~occ_categ+sex, scales = "free") + 
  scale_color_viridis_d() +  
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1))
```

## Decompose by skill tercile

```{r}
acs[average_value_skills >= .66, skill_tercile := "High Skill"]
acs[average_value_skills < .66 &average_value_skills >= .33, skill_tercile := "Medium Skill"]
acs[average_value_skills < .33, skill_tercile := "Low Skill"]

acs[average_value_skills > .5, skill_half := "High Skill"]
acs[average_value_skills <= .5, skill_half := "Low Skill"]

# 
acs[, occsoc_sub := substr(OCCSOC,1,4)]
# see if I can recreate sakamoto's graphs
r_sq_dt <- data.table()
i <- 0
for(c.year in unique(acs$year)){
  for(c.skill in unique(acs[!is.na(skill_tercile)]$skill_tercile)){
    i <- i + 1
    #print(i)
    out <- lm(log_incwage ~ occsoc_sub, data = acs[year == c.year & skill_tercile == c.skill])
    out_dt <- data.table(year = c.year, skill_tercile = c.skill,
                         r_sq = summary(out)$r.squared)
    r_sq_dt <- rbind(r_sq_dt, out_dt, fill = T)
  }
}

ggplot(r_sq_dt) + 
  geom_point(aes(x = year, y = r_sq, color = skill_tercile)) + 
  geom_line(aes(x = year, y = r_sq, color = skill_tercile, group = skill_tercile)) + 
  labs(title = "R-Squared for Occupation (114 categories)\nRegressed on Log(Income)")+
    theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1))



r_sq_dt <- data.table()
i <- 0
for(c.year in unique(acs$year)){
  for(c.skill in unique(acs[!is.na(skill_half)]$skill_half)){
    for(c.sex in list("male", "female", c("male", "female"))){
      i <- i + 1
      #print(i)
      out <- lm(log_incwage ~ OCCSOC, data = acs[year == c.year & skill_half == c.skill & 
                                                       sex %in% c.sex])
      out_dt <- data.table(year = c.year, skill_tercile = c.skill,
                           sex = paste0(c.sex, collapse = ","),
                           r_sq = summary(out)$r.squared)
      r_sq_dt <- rbind(r_sq_dt, out_dt, fill = T)
    }
  }
}

ggplot(r_sq_dt) + 
  geom_point(aes(x = year, y = r_sq, color = skill_tercile)) + 
  geom_line(aes(x = year, y = r_sq, color = skill_tercile, group = skill_tercile)) + 
  labs(title = "R-Squared for Occupation (Most Detailed)\nRegressed on Log(Income)") + 
  facet_wrap(~sex)


```

## Does inequality in skills track inequality in earnings?

### Is there a return to education re: skills

```{r}
acs[, ed_num := as.numeric(as.character(factor(educ, labels = c(0,2.5, 6.5, 9,10,11,12, 13,14,16,18))))]
acs[ed_num <= 11, ed_categ := "Less than High School"]
acs[ed_num > 11 &ed_num < 16, ed_categ := "HS or Some College"]
acs[ed_num >= 16, ed_categ := "College Plus"]

skills_overview4 <- acs[,.(pc1 = weighted.mean(pc1,w = perwt, na.rm = T),
                           pc2 = weighted.mean(pc2,w = perwt,  na.rm = T), 
                           pc3 = weighted.mean(pc3,w = perwt,  na.rm = T),
                           pc4 = weighted.mean(pc4,w = perwt,  na.rm = T),
                           programming = weighted.mean(programming,w = perwt,  na.rm = T),
                           tech_skills = weighted.mean(tech_skills, w = perwt, na.rm = T), average_value_skills = weighted.mean(average_value_skills, w = perwt, na.rm = T)), by = .(year, age_cat, sex, ed_categ)]

skills_overview4_melt <- melt(skills_overview4, id.vars = c("year", "age_cat", "sex", "ed_categ"))

skills_overview4_melt[, cohort := floor((as.numeric(year) - as.numeric(substr(age_cat,1,2)))/10)*10]

skills_overview4_melt <- skills_overview4_melt[,.(value = mean(value)), by = .(cohort, ed_categ,
                                                                               variable, year, sex)]

ggplot(skills_overview4_melt[variable %like% "pc1|progr|tech|averag"]) + 
  geom_line(aes(x = year, y = value, color = as.factor(cohort), group = as.factor(cohort))) + 
  facet_grid( variable + sex~ed_categ, scales = "free") + 
  scale_color_viridis_d()+  
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1))
```

# Examine how class of worker, industry, and ed affect inequality (switch to censuses)

```{r, echo = F}
rm(list = ls())

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(ggplot2)
library(dplyr)
library(magrittr)
library(stringr)
library(readxl)
library(ggrepel)
library(foreign)
library(ipumsr)
library(data.table)



options
theme_set(theme_bw())

# load the data
setwd("/Users/hyork/Documents/projects/occupation/code")

weighted.var <- function(x, w, na.rm = FALSE) {
  if (na.rm) {
    w <- w[i <- !is.na(x)]
    x <- x[i]
  }
  sum.w <- sum(w)
  sum.w2 <- sum(w^2)
  mean.w <- sum(x * w) / sum(w)
  (sum.w / (sum.w^2 - sum.w2)) * sum(w * (x - mean.w)^2, na.rm =
                                       na.rm)
}


# #census_1940 <- readRDS("../inputs/usa_00005.rds")
# census_1940 <- readstata13::read.dta13("../inputs/usa_00009.dta")
# census_1940 <- data.table(census_1940)
# #keep vars
# census_1940 <- census_1940[,.(race, hispan, year,statefip, countyicp,metro,perwt, sex, age, school,educ, empstat,occ,occ1950, ind,ind1950, classwkr, wkswork2, incwage)]
# 
# # subset
# census_1940 <- census_1940[empstat == "employed" &
#                              (school == "no, not in school"| as.numeric(as.character(age)) >=34) &
#                              as.numeric(as.character(age)) %in% 25:64 &
#                             wkswork2 %like% "40|48|50" &
#                              incwage != 999998 &
#                              incwage > 0]

# #saveRDS(census_1940, "../inputs/usa_00009.rds")
census_1940 <- readRDS( "../inputs/usa_00009.rds")

setnames(census_1940, c("occ1950", "ind1950"), c("origocc1950", "origind1950"))
# derive a 5-year age var
census_1940[, age_cat := paste0(floor(as.numeric(as.character(age))/5)*5,
                                " to ", 
                                floor(as.numeric(as.character(age))/5)*5 + 4)]

# make occ var char
census_1940[, occ := as.character(occ)]

# 
census_1940[nchar(occ) == 1, occ := paste0("00", occ)]
census_1940[nchar(occ) == 2, occ := paste0("0", occ)]

# make first dig var
census_1940[, first_dig := substr(occ, 1,1)]

# make map for first dig categories
map <- data.table(first_dig = as.character(0:9), 
                  occ_categ = c("Professional", "Clerical", "Service",
                                "Agricultural, etc.", "Skilled", "Skilled",
                                "Semiskilled", "Semiskilled", "Unskilled",
                                "Unskilled"))
census_1940 <- merge(census_1940, map, by = "first_dig")

# merge census 1940 file on ind codes
ind_codes <- read_xlsx("../ref/IND1940.xlsx") %>% data.table()
ind_codes[, IND := as.numeric(IND)]
census_1940 <- merge(census_1940, 
                     ind_codes[!is.na(IND) & IND != 996], 
                     by.x = "ind", by.y = "IND", all.x = T)

# crosswalk to ppp
ppp <- read_excel("../ref/CPI_U_RS.xlsx") %>% data.table
ppp[1, year := 1940]
ppp[1, cpi := 32]
ppp[, year := as.character(year)]
ppp[, xwalk_fac := cpi/369.8]
census_1940 <- merge(census_1940, ppp, by = "year")
census_1940[, incwage := incwage/xwalk_fac]
# create log inc wage
census_1940[, log_incwage := log(incwage + 1)]
```

```{r}
## track change in ed diversity across time compared to income diversity
census_1940[, ed_num := as.numeric(as.character(factor(educ, labels = c(0,2.5, 6.5, 9,10,11,12, 13,14,15,16,18))))]
inc_ed_ineq <- census_1940[,.(ed_gini = DescTools::Gini(ed_num),
                              income_gini = DescTools::Gini(incwage), 
                              N = .N), by = .(occ, year, occ_categ, sex)]

ggplot(inc_ed_ineq[N > 50]) + 
  geom_point(aes(x = ed_gini, y = income_gini, color = occ_categ), size = .5, shape = 1) + 
  facet_grid(year~sex) + 
  xlim(0, .5) + 
  ylim(0, .5) +  theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1))

```

```{r}
## track change in ed diversity across time compared to income diversity
census_1940[, ed_num := as.numeric(as.character(factor(educ, labels = c(0,2.5, 6.5, 9,10,11,12, 13,14,15,16,18))))]
inc_ed_ineq <- census_1940[,.(ed_gini = DescTools::Gini(ed_num),
                              income_gini = DescTools::Gini(incwage), 
                              N = .N), by = .(origocc1950, year, occ_categ, sex)]

ggplot(inc_ed_ineq[N > 50]) + 
  geom_point(aes(x = ed_gini, y = income_gini, color = occ_categ), size = .5, shape = 1) + 
  facet_grid(year~sex) + 
  xlim(0, .5) + 
  ylim(0, .5)

census_1940[,N_occ_ed := .N, by = .(origocc1950, educ, year, sex)]
inc_ed_ineq_2 <- census_1940[!is.na(log_incwage) &N_occ_ed > 1,.(within_ed_occ_var = var(log_incwage),
                                                                 N_ed_occ = .N,
                                                                 ed_occ_avg = mean(log_incwage)), by = .(origocc1950, occ_categ, educ, year, sex)] %>% 
  merge(., census_1940[!is.na(log_incwage)&N_occ_ed > 1,.(witihin_occ_var = var(log_incwage),
                                                          N_occ = .N,
                                                          occ_avg = mean(log_incwage)), by = .(origocc1950, year, sex)], by = c("origocc1950", "year", "sex"))

inc_ed_ineq_2_sum  <- inc_ed_ineq_2[,.(bw_var = weighted.var(ed_occ_avg, N_ed_occ),
                                       wi_var = mean(witihin_occ_var),
                                       occ_avg = mean(occ_avg),
                                       N_occ = mean(N_occ)), by = .(origocc1950, year, sex)]

inc_ed_ineq_2_sum[, bw_perc := bw_var /(wi_var)]

census_1940[year == 2018,.(origocc1950, occ_categ)] %>% unique() %>% 
  merge(., inc_ed_ineq_2_sum, by = "origocc1950") -> inc_ed_ineq_2_sum

ggplot(inc_ed_ineq_2_sum[N_occ > 50 & !origocc1950 %like% "nec|n.e.c.|NEC|missing|unknown|NA"]) + 
  geom_line(aes(x = as.numeric(year), y = bw_perc, group = (origocc1950)), size = .5, alpha = .25)+
  facet_grid(sex~occ_categ) + 
  scale_color_viridis_d() +
  guides(color = F) +
  ylim(0, 1) +
  scale_y_continuous(trans = "log10")+
  labs(title = "Between-Educational Category,\nWithin-Occupation Variance(log_income)",
       x = "Year",
       y = "Between-Education Variance as a Percent of \nTotal Variance Within Each Occupation") +
    theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1))


```

## Repeat for Industry

```{r}
census_1940[,N_occ_ind:= .N, by = .(origocc1950, origind1950, year, sex)]
inc_ind_ineq_2 <- census_1940[!is.na(log_incwage) &N_occ_ind > 1,.(within_ind_occ_var = var(log_incwage),
                                                                   N_ind_occ = .N,
                                                                   ind_occ_avg = mean(log_incwage)), by = .(origocc1950, occ_categ, origind1950, year, sex)] %>% 
  merge(., census_1940[!is.na(log_incwage)&N_occ_ind > 1,.(witihin_occ_var = var(log_incwage),
                                                           N_occ = .N,
                                                           occ_avg = mean(log_incwage)), by = .(origocc1950, year, sex)], by = c("origocc1950", "year", "sex"))

inc_ind_ineq_2_sum  <- inc_ind_ineq_2[,.(bw_var = weighted.var(ind_occ_avg, N_ind_occ),
                                         wi_var = mean(witihin_occ_var),
                                         occ_avg = mean(occ_avg),
                                         N_occ = mean(N_occ)), by = .(origocc1950, year, sex)]

inc_ind_ineq_2_sum[, bw_perc := bw_var /(wi_var)]

census_1940[year == 2018,.(origocc1950, occ_categ)] %>% unique() %>% 
  merge(., inc_ind_ineq_2_sum, by = "origocc1950") -> inc_ind_ineq_2_sum

ggplot(inc_ind_ineq_2_sum[N_occ > 50 & !origocc1950 %like% "nec|n.e.c.|NEC|missing|unknown|NA"]) + 
  geom_line(aes(x = as.numeric(year), y = bw_perc, group = (origocc1950)), size = .5, alpha = .25)+
  facet_grid(sex~occ_categ) + 
  scale_color_viridis_d() +
  guides(color = F) +
  ylim(0, 1) +
  scale_y_continuous(trans = "log10")+
  labs(title = "Between-Industry Category,\nWithin-Occupation Variance(log_income)",
       x = "Year",
       y = "Between-Industry Variance as a Percent of \nTotal Variance Within Each Occupation")+
    theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1))


```

## Repeat for Industry

```{r}
census_1940[,N_occ_ind:= .N, by = .(origocc1950, origind1950, year, sex)]
inc_ind_ineq_2 <- census_1940[!is.na(log_incwage) &N_occ_ind > 1,.(within_ind_occ_var = var(log_incwage),
                                                                   N_ind_occ = .N,
                                                                   ind_occ_avg = mean(log_incwage)), by = .(origocc1950, occ_categ, origind1950, year, sex)] %>% 
  merge(., census_1940[!is.na(log_incwage)&N_occ_ind > 1,.(witihin_occ_var = var(log_incwage),
                                                           N_occ = .N,
                                                           occ_avg = mean(log_incwage)), by = .(origind1950, year, sex)], by = c("origind1950", "year", "sex"))

inc_ind_ineq_2_sum  <- inc_ind_ineq_2[,.(bw_var = weighted.var(ind_occ_avg, N_ind_occ),
                                         wi_var = mean(witihin_occ_var),
                                         occ_avg = mean(occ_avg),
                                         N_occ = mean(N_occ)), by = .(origind1950, year, sex)]

inc_ind_ineq_2_sum[, bw_perc := bw_var /(wi_var)]


ggplot(inc_ind_ineq_2_sum[N_occ > 50 & !origind1950 %like% "nec|n.e.c.|NEC|missing|unknown|NA"]) + 
  geom_line(aes(x = as.numeric(year), y = bw_perc, group = (origind1950)), size = .5, alpha = .25)+
  facet_grid(sex~.) + 
  scale_color_viridis_d() +
  guides(color = F) +
  ylim(0, 1) +
  scale_y_continuous(trans = "log10")+
  labs(title = "Between-Occupation Category,\nWithin-Industry Variance(log_income)",
       x = "Year",
       y = "Between-Occupation Variance as a Percent of \nTotal Variance Within Each Industry")+
    theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1))


```


## Repeat for Industry

```{r}
census_1940[,N_occ_ind:= .N, by = .(origocc1950, origind1950, year, sex)]
inc_ind_ineq_2 <- census_1940[!is.na(log_incwage) &N_occ_ind > 1,.(within_ind_occ_var = var(log_incwage),
                                                                   N_ind_occ = .N,
                                                                   ind_occ_avg = mean(log_incwage)), by = .(origocc1950, occ_categ, origind1950, year, sex)] %>% 
  merge(., census_1940[!is.na(log_incwage)&N_occ_ind > 1,.(witihin_occ_var = var(log_incwage),
                                                           N_occ = .N,
                                                           occ_avg = mean(log_incwage)), by = .(origocc1950, year, sex)], by = c("origocc1950", "year", "sex"))

inc_ind_ineq_2_sum  <- inc_ind_ineq_2[,.(bw_var = weighted.var(ind_occ_avg, N_ind_occ),
                                         wi_var = mean(witihin_occ_var),
                                         occ_avg = mean(occ_avg),
                                         N_occ = mean(N_occ)), by = .(origocc1950, year, sex)]

inc_ind_ineq_2_sum[, bw_perc := bw_var /(wi_var)]

census_1940[year == 2018,.(origocc1950, occ_categ)] %>% unique() %>% 
  merge(., inc_ind_ineq_2_sum, by = "origocc1950") -> inc_ind_ineq_2_sum

ggplot(inc_ind_ineq_2_sum[N_occ > 50 & !origocc1950 %like% "nec|n.e.c.|NEC|missing|unknown|NA"]) + 
  geom_line(aes(x = as.numeric(year), y = bw_perc, group = (origocc1950)), size = .5, alpha = .25)+
  facet_grid(sex~occ_categ) + 
  scale_color_viridis_d() +
  guides(color = F) +
  ylim(0, 1) +
  scale_y_continuous(trans = "log10")+
  labs(title = "Between-Industry Category,\nWithin-Occupation Variance(log_income)",
       x = "Year",
       y = "Between-Industry Variance as a Percent of \nTotal Variance Within Each Occupation")+
    theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1))


```

## Repeat within demographic group

```{r}
census_1940[ed_num <= 11, ed_categ := "Less than High School"]
census_1940[ed_num > 11 &ed_num < 16, ed_categ := "HS or Some College"]
census_1940[ed_num >= 16, ed_categ := "College Plus"]

census_1940[race %like% "black|white", demographic := paste(race, ed_categ, sep = "\n")]
census_1940[,N_occ_ind:= .N, by = .(origocc1950, demographic, year, sex)]
inc_ind_ineq_2 <- census_1940[!is.na(log_incwage) &N_occ_ind > 1,.(within_ind_occ_var = var(log_incwage),
                                                                   N_ind_occ = .N,
                                                                   ind_occ_avg = mean(log_incwage)), by = .(origocc1950, occ_categ, demographic, year, sex)] %>% 
  merge(., census_1940[!is.na(log_incwage)&N_occ_ind > 1,.(witihin_occ_var = var(log_incwage),
                                                           N_occ = .N,
                                                           occ_avg = mean(log_incwage)), by = .(demographic, year, sex)], by = c("demographic", "year", "sex"))

inc_ind_ineq_2_sum  <- inc_ind_ineq_2[,.(bw_var = weighted.var(ind_occ_avg, N_ind_occ),
                                         wi_var = mean(witihin_occ_var),
                                         occ_avg = mean(occ_avg),
                                         N_occ = mean(N_occ)), by = .(demographic, year, sex)]

inc_ind_ineq_2_sum[, bw_perc := bw_var /(wi_var)]


ggplot(inc_ind_ineq_2_sum[N_occ > 50]) + 
  geom_line(aes(x = as.numeric(year), y = bw_perc, group = (demographic),
                color = demographic), size = 1)+
  facet_grid(sex~.) + 
  scale_color_viridis_d() +
  ylim(0, 1) +
  scale_y_continuous(trans = "log10")+
  labs(title = "Between-Occupation Category,\nWithin-Demographic Group Variance(log_income)",
       x = "Year",
       y = "Between-Occupation Variance as a Percent of \nTotal Variance Within Each Demogrpahic Group")+
    theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1))


```