---
title: "analysis_v"
author: "Hunter York"
date: "11/18/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(ggplot2)
library(dplyr)
library(magrittr)
library(stringr)
library(readxl)
library(ggrepel)
library(foreign)
library(ipumsr)
library(data.table)
library(factoextra)



options
theme_set(theme_bw())

# load the data
setwd("/Users/hyork/Documents/projects/occupation/code")

weighted.var <- function(x, w, na.rm = FALSE) {
  if (na.rm) {
    w <- w[i <- !is.na(x)]
    x <- x[i]
  }
  sum.w <- sum(w)
  sum.w2 <- sum(w^2)
  mean.w <- sum(x * w) / sum(w)
  (sum.w / (sum.w^2 - sum.w2)) * sum(w * (x - mean.w)^2, na.rm =
                                       na.rm)
}



# acs <- readstata13::read.dta13("../inputs/cps_00004.dta")
# acs2 <- fread("../inputs/cps_00005.csv")
# acs[,-tolower(c("OCC2010", "OCC1990", "OCC", "OCCLY", "OCC90LY", "OCC10LY"))]
# acs <- cbind(acs, acs2[,.(OCC2010, OCC1990, OCC, OCCLY, OCC90LY, OCC10LY)])
# acs <- acs[acs$empstat == "at work" &
#                              (acs$schlcoll %like% "does not attend|niu"| is.na(acs$schlcoll)) &
#                              as.numeric(as.character(acs$age)) %in% 25:64 &
#                             acs$wkswork1 >= 30 &
#                              acs$incwage != 99999999 &
#                              acs$incwage > 0,]
# acs <- data.table(acs)
# #keep vars
# acs <- acs[,.(race, hispan, year,asecwt, sex, age, schlcoll,educ, empstat,OCC,OCC2010,OCC1990, ind,ind1990, classwkr, ahrsworkt,wkswork1, uhrsworkly,classwly,workly,OCCLY,indly,OCC90LY,ind90ly,OCC10LY, incwage)]

# subset


#saveRDS(acs, "../inputs/cps_00004.rds")
acs <- readRDS( "../inputs/cps_00004.rds")

# derive a 5-year age var
acs[, age_cat := paste0(floor(as.numeric(as.character(age))/5)*5,
                        " to ", 
                        floor(as.numeric(as.character(age))/5)*5 + 4)]

# make first dig var
#acs[, first_dig := substr(occ, 1,1)]

# make map for first dig categories
# map <- data.table(first_dig = as.character(0:9), 
#                   occ_categ = c("Professional", "Clerical", "Service",
#                                 "Agricultural, etc.", "Skilled", "Skilled",
#                                 "Semiskilled", "Semiskilled", "Unskilled",
#                                 "Unskilled"))
# acs <- merge(acs, map, by = "first_dig")

# # crosswalk to ppp
# ppp <- read_excel("../ref/CPI_U_RS.xlsx") %>% data.table
# ppp[1, year := 1940]
# ppp[1, cpi := 33]
# ppp[, year := as.character(year)]
# ppp[, xwalk_fac := cpi/369.8]
# acs <- merge(acs, ppp, by = "year")
# acs[, incwage := incwage/xwalk_fac]
# create log inc wage
acs[, log_incwage := log(incwage + 1)]

# load occ_soc xwalk
occ_xwalk <- data.table(read_excel("../ref/nem-occcode-cps-crosswalk.xlsx"))
names(occ_xwalk) <- occ_xwalk[4,] %>% unlist()
occ_xwalk <- occ_xwalk[-(1:4),]
occ_xwalk[, OCCSOC := gsub("-", "", `Hybrid SOC Code`)]

# now create skills dataset
skills_2009 <- read.delim('../ref/db_14_0 2009.7/Skills.txt') %>% data.table()
skills_2013 <-  read.delim('../ref/db_18_0_2013.7/Skills.txt') %>% data.table()
skills_2018 <- read_excel("../ref/db_22_2_excel 2018.2/Skills.xlsx") %>% data.table()
skills_2009[, year := 2009]
skills_2013[, year := 2013]
skills_2018[, year := 2018]
setnames(skills_2018, names(skills_2018), gsub(" ", ".", names(skills_2018), fixed = T))
skills_2018[, `O.NET.SOC.Code` := `O*NET-SOC.Code`]
skills <- rbindlist(list(skills_2009, skills_2013, skills_2018), fill = T)
skills <- skills[, .(O.NET.SOC.Code, Element.Name, Scale.ID, Data.Value, Standard.Error, year)]

# reformate onet codes to merge
skills[, OCCSOC := gsub("-", "", substr(O.NET.SOC.Code,1,7))]
skills <- skills[Scale.ID == "LV",.(Data.Value = mean(Data.Value),
                                    Standard.Error = mean(as.numeric(Standard.Error))), by = .(Element.Name,Scale.ID, OCCSOC)]

# chack to see which occ codes are missing
occ_xwalk[!OCCSOC %in% unique(skills$OCCSOC), unique(OCCSOC)]

# 
skills[,Data.Value := percent_rank(Data.Value), by = .(Element.Name, Scale.ID)]
skills_wide <- dcast(skills[Scale.ID == "LV"], OCCSOC   ~Element.Name, value.var = "Data.Value")
res.pca <- prcomp(skills_wide[,2:34], scale = TRUE, center = T)
# skills wide
skills_sum <- skills_wide
skills_sum[, pc1 := predict(res.pca, newdata = .SD)[,1], .SDcols = names(skills_sum)]
skills_sum[, pc2 := predict(res.pca, newdata = .SD)[,2], .SDcols = names(skills_sum)]
skills_sum[, pc3 := predict(res.pca, newdata = .SD)[,3], .SDcols = names(skills_sum)]
skills_sum[, pc4 := predict(res.pca, newdata = .SD)[,4], .SDcols = names(skills_sum)]

skills_sum[, programming := Programming]
skills_sum[, tech_skills := Programming + `Complex Problem Solving` +
             `Mathematics` + Programming + Science + `Systems Analysis` + 
             Troubleshooting]
skills_sum[, average_value_skills := rowMeans(.SD), .SDcols = rownames(res.pca$rotation)]
skills_sum <- merge(skills_sum, occ_xwalk, by = "OCCSOC")

# collapse to OCCSOC
vars <- names(skills_wide)[-1]
skills_final <- skills_sum[,lapply(.SD, mean), .SDcols = vars, by = .(`CPS Code`, `CPS Occupational Title`)]
acs[, OCC2010 := as.character(OCC2010)]

# merge it all
acs <- merge(acs, skills_final, 
             all.x = T,
             by.x = "OCC2010",
             by.y = "CPS Code")

# 


```

# Compositional Changes

## Plot top 35 Occupations change in % share over time

```{r}
occ_share <- acs[!is.na(`CPS Occupational Title`),.(N = .N), by = .(`CPS Occupational Title`, OCC2010, year, sex)]
occ_share[,percent_share := N/sum(N), by = .(year, sex)]
occ_share[, year_rank := frankv(percent_share, order = -1L), by = .(year, sex)]
occ_share <- occ_share[`CPS Occupational Title` %in% occ_share[year %in% c(1981, 2020) & year_rank <= 20, unique(`CPS Occupational Title`)]]
occ_share[, `CPS Occupational Title` := paste0(str_sub(`CPS Occupational Title`,1,10), "...\n", 
                                               str_sub(`CPS Occupational Title`,-10,-1))]
occ_share[, `CPS Occupational Title` := factor(`CPS Occupational Title`, levels = occ_share[year == 2020 & sex == "male"] %>% .[order(percent_share, decreasing = T), `CPS Occupational Title`])]
ggplot(occ_share) + 
  geom_line(aes(x = year, y = percent_share*100, group = paste(`CPS Occupational Title`, sex), color = sex)) + 
  facet_wrap(~`CPS Occupational Title`)+
    theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1))+
    scale_y_continuous(trans = "log10")+
  theme(strip.text = element_text(size = 5), axis.text = element_text(size = 5))


```

## Repeat for younger ages (25-34)

```{r}
occ_share <- acs[!is.na(`CPS Occupational Title`) & as.numeric(as.character(age)) <= 34,.(N = .N), by = .(`CPS Occupational Title`, OCC2010, year, sex)]
occ_share[,percent_share := N/sum(N), by = .(year, sex)]
occ_share[, year_rank := frankv(percent_share, order = -1L), by = .(year, sex)]
occ_share <- occ_share[`CPS Occupational Title` %in% occ_share[year %in% c(1981, 2020) & year_rank <= 20, unique(`CPS Occupational Title`)]]
occ_share[, `CPS Occupational Title` := paste0(str_sub(`CPS Occupational Title`,1,10), "...\n", 
                                               str_sub(`CPS Occupational Title`,-10,-1))]
occ_share[, `CPS Occupational Title` := factor(`CPS Occupational Title`, levels = occ_share[year == 2020 & sex == "male"] %>% .[order(percent_share, decreasing = T), `CPS Occupational Title`])]
ggplot(occ_share) + 
  geom_line(aes(x = year, y = percent_share*100, group = paste(`CPS Occupational Title`, sex), color = sex)) + 
  facet_wrap(~`CPS Occupational Title`)+
    theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1)) +
  scale_y_continuous(trans = "log10") + 
  theme(strip.text = element_text(size = 5), axis.text = element_text(size = 5))

```

## Plot average value for each skill over time, all ages

```{r}
skill_share <- acs[, lapply(.SD, weighted.mean, w = asecwt, na.rm = T), .SDcols = vars, by = .(sex, year)]
skill_share <- melt(skill_share, id.vars = c("year", "sex"))
skill_share[, variable := substr(variable, 1,15)]
skill_share[, variable := factor(variable, levels = skill_share[year == 1981 & sex == "male"] %>% .[order(value, decreasing = T), variable])]
ggplot(skill_share[!variable %like% "pc|average|tech"]) + 
  geom_line(aes(x = year, y = value, group = paste0(sex, variable), color = sex))+
facet_wrap(~variable) +
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1)) + 
  theme(strip.text = element_text(size = 5), axis.text = element_text(size = 5))
```
## Repeat for younger ages (25-34)

```{r}
skill_share <- acs[as.numeric(as.character(age)) <= 34, lapply(.SD, weighted.mean, w = asecwt, na.rm = T), .SDcols = vars, by = .(sex, year)]
skill_share <- melt(skill_share, id.vars = c("year", "sex"))
skill_share[, variable := substr(variable, 1,15)]
skill_share[, variable := factor(variable, levels = skill_share[year == 1981 & sex == "male"] %>% .[order(value, decreasing = T), variable])]
ggplot(skill_share[!variable %like% "pc|average|tech"]) + 
  geom_line(aes(x = year, y = value, group = paste0(sex, variable), color = sex))+
facet_wrap(~variable) +
  theme(axis.text.x = element_text(angle = -90, vjust = 0.5, hjust=1)) + 
  theme(strip.text = element_text(size = 5), axis.text = element_text(size = 5))
```


# See if movement between occupations clusters by skill, at the cohort level

## First, visualize movement between occupaitons

```{r}
acs[as.numeric(as.character(age)) <= 59, cohort := floor((year - as.numeric(as.character(age)))/10)*10]

ggplot(acs) + 
  geom_density(aes(x = average_value_skills, color = age_cat, fill = age_cat), alpha = .2) + 
  facet_wrap(~cohort) + 
  ggtitle("Average Value Skills")

ggplot(acs) + 
  geom_density(aes(x = tech_skills, color = age_cat, fill = age_cat), alpha = .2) + 
  facet_wrap(~cohort) + 
  ggtitle("Tech Skills")

ggplot(acs) + 
  geom_density(aes(x = programming, color = age_cat, fill = age_cat), alpha = .2) + 
  facet_wrap(~cohort) + 
  ggtitle("Programming Skills")

ggplot(acs) + 
  geom_density(aes(x = pc1, color = age_cat, fill = age_cat), alpha = .2) + 
  facet_wrap(~cohort) + 
  ggtitle("PC1")

ggplot(acs) + 
  geom_density(aes(x = pc2, color = age_cat, fill = age_cat), alpha = .2) + 
  facet_wrap(~cohort) + 
  ggtitle("PC2")

ggplot(acs) + 
  geom_density(aes(x = pc3, color = age_cat, fill = age_cat), alpha = .2) + 
  facet_wrap(~cohort) + 
  ggtitle("PC3")

ggplot(acs) + 
  geom_density(aes(x = pc4, color = age_cat, fill = age_cat), alpha = .1) + 
  facet_wrap(~cohort) + 
  ggtitle("PC4")
```

# See if movement between occupations clusters by skill, at the individual level

## First, visualize movement between occupaitons

```{r}
# merge on both new and old jobs
setnames(acs, vars, paste0(vars, "_current"))
setnames(acs, "CPS Occupational Title","CPS Occupational Title_current")
# merge it all
acs[, OCC10LY := as.character(OCC10LY)]
acs <- merge(acs, skills_final, 
             all.x = T,
             by.x = "OCC10LY",
             by.y = "CPS Code")

setnames(acs, vars, paste0(vars, "_ly"))
setnames(acs, "CPS Occupational Title","CPS Occupational Title_ly")

# subset to places where people have moved jobs
acs_moved <- acs[OCC2010 != OCC10LY]

# calculate flows
acs_flows <- acs[!is.na(`CPS Occupational Title_current`) & 
                   !is.na(`CPS Occupational Title_ly`),.(mvmt = .N), by = .(OCC2010, OCC10LY, `CPS Occupational Title_current`, `CPS Occupational Title_ly`)]

# graph top movment
temp <- acs_moved[, .N, by = OCC10LY] %>% .[order(N, decreasing = T)] %>% .[1:7, OCC10LY]
acs_flows[,rank := frankv(mvmt, order = -1L, ties.method = "first"), by = OCC10LY]
acs_flows[, OCC_title_ly := paste0(str_sub(`CPS Occupational Title_ly`, 1, 15), 
                                   "...")]
acs_flows[, OCC_title_current:= paste0(str_sub(`CPS Occupational Title_current`, 1, 15), 
                                   "...")]

acs_flows[, OCC_title_current := factor(OCC_title_current)]

ggplot(acs_flows[OCC10LY %in% temp & rank %in% 2:6]) +
  geom_bar(aes(x = OCC_title_ly, y = mvmt, fill = OCC_title_current), position = position_stack(),
           stat = "identity") + 
  geom_text(aes(x = OCC_title_ly, y = mvmt,label = OCC_title_current, group = OCC_title_current),
             position = position_stack(vjust = 0.5)) + 
  guides(fill = F) + 
  theme(strip.text = element_text(size = 5), axis.text = element_text(size = 5))
```

## See how these cluster by skill

```{r, out.width="600px"}
skills_subs <- acs[!is.na(`CPS Occupational Title_current`) &
                     !is.na(`CPS Occupational Title_ly`),.SD, .SDcols = names(acs)[names(acs) %like% "_ly|_current"]] %>% unique()


acs_flows_merged <- merge(acs_flows, skills_subs)

acs_flows_merged_melt <- melt(acs_flows_merged, id.vars = c("CPS Occupational Title_current", 
                                                            "CPS Occupational Title_ly",
                                                            "OCC2010", "OCC10LY",
                                                            "mvmt", "rank"))

acs_flows_merged_melt[, value := as.numeric(value)]
acs_flows_merged_melt <- acs_flows_merged_melt[!is.na(value)]

# cast wide
acs_flows_merged_melt[, current := ifelse(variable %like% "current", "current", "ly")]
acs_flows_merged_melt[, variable := gsub("_current|_ly", "", variable)]
acs_flows_merged_wide <- dcast(acs_flows_merged_melt, ... ~ current, value.var = "value")

# rate standardize
acs_flows_merged_wide[, mvmt_percent := mvmt/sum(mvmt), by = .(OCC10LY, variable)]
# plot
ggplot(acs_flows_merged_wide[OCC2010 != OCC10LY & 
                               mvmt > 20 & variable %in% unique(acs_flows_merged_wide$variable)[1:21]]) +
  geom_point(aes(x = current, y = ly, size = mvmt_percent, alpha = mvmt)) + 
  facet_wrap(~variable, scales = "free") + 
  scale_alpha_continuous(trans = "log10") + 
  scale_size_continuous(range = c(0, 2)) + 
  theme(strip.text = element_text(size = 5), axis.text = element_text(size = 5))

ggplot(acs_flows_merged_wide[OCC2010 != OCC10LY & 
                               mvmt > 20 & variable %in% unique(acs_flows_merged_wide$variable)[22:42]]) +
  geom_point(aes(x = current, y = ly, size = mvmt_percent, alpha = mvmt)) + 
  facet_wrap(~variable, scales = "free") + 
  scale_alpha_continuous(trans = "log10")  + 
  scale_size_continuous(range = c(0, 2)) + 
  theme(strip.text = element_text(size = 5), axis.text = element_text(size = 5))
```

# See if we can recreate the sakomoto graphs from last week

```{r}
acs2 <- data.table(readstata13::read.dta13("../inputs/cps_00004.dta"))
acs2 <- acs2[year %in% seq(1981, 2020, 6)]
#acs2 <- acs2[year %in% c(1981,2020)]

#acs2 <- acs2[
  # acs2$empstat == "at work" &
  #                            (acs2$schlcoll %like% "does not attend|niu"| is.na(acs2$schlcoll)) &
  #                            as.numeric(as.character(acs2$age)) %in% 25:64 &
                             #acs2$wkswork1 >= 30 & ### !!!!! ahhhh
                             #acs2$incwage != 99999999,]
                            # acs2$incwage > 0,]
acs2[, log_incwage := log(incwage + 1)]
acs2[, ed_num := as.numeric(as.character(factor(educ,levels = levels(acs2$educ), labels = c(0,0,0,2.5,1,2,3,4,5.5,
                                                                5,6,7.5, 7,8,9,10,11,11, 11, 11,12,
                                                                13,14,14,15,15,15,16,16,17,17,18,18,18,18,NA))))]

acs2[ed_num <= 14, ed_categ := "Less than HS"]
acs2[ed_num > 14, ed_categ := "College Plus"]

# r_sq_dt <- data.table()
# i <- 0
# for(c.year in unique(acs2$year)){
#   for(c.educ_yr in unique(acs2[!is.na(ed_categ)]$ed_categ)){
#     for(c.sex in list("male", "female", c("male", "female"))){
#       i <- i + 1
#       print(i)
#       out <- lm(log_incwage ~ as.factor(occ2010), data = acs2[year == c.year & ed_categ == c.educ_yr & 
#                                                        sex %in% c.sex])
#       out_dt <- data.table(year = c.year, ed_categ = c.educ_yr,
#                            sex = paste0(c.sex, collapse = ","),
#                            r_sq = summary(out)$r.squared)
#       r_sq_dt <- rbind(r_sq_dt, out_dt, fill = T)
#     }
#   }
# }
# 
# ggplot(r_sq_dt) + 
#   geom_point(aes(x = year, y = r_sq, color = ed_categ)) + 
#   geom_line(aes(x = year, y = r_sq, color = ed_categ, group = ed_categ)) + 
#   labs(title = "R-Squared for Occupation (Most Detailed)\nRegressed on Log(Income), all Cases") + 
#   facet_wrap(~sex)

```

### Drop missing

```{r}

acs2 <- acs2[
  # acs2$empstat == "at work" &
  #                            (acs2$schlcoll %like% "does not attend|niu"| is.na(acs2$schlcoll)) &
  #                            as.numeric(as.character(acs2$age)) %in% 25:64 &
                             #acs2$wkswork1 >= 30 & ### !!!!! ahhhh
                             acs2$incwage != 99999999,]
                            # acs2$incwage > 0,]

# r_sq_dt <- data.table()
# i <- 0
# for(c.year in unique(acs2$year)){
#   for(c.educ_yr in unique(acs2[!is.na(ed_categ)]$ed_categ)){
#     for(c.sex in list("male", "female", c("male", "female"))){
#       i <- i + 1
#       print(i)
#       out <- lm(log_incwage ~ as.factor(occ2010), data = acs2[year == c.year & ed_categ == c.educ_yr & 
#                                                        sex %in% c.sex])
#       out_dt <- data.table(year = c.year, ed_categ = c.educ_yr,
#                            sex = paste0(c.sex, collapse = ","),
#                            r_sq = summary(out)$r.squared)
#       r_sq_dt <- rbind(r_sq_dt, out_dt, fill = T)
#     }
#   }
# }
# 
# ggplot(r_sq_dt) + 
#   geom_point(aes(x = year, y = r_sq, color = ed_categ)) + 
#   geom_line(aes(x = year, y = r_sq, color = ed_categ, group = ed_categ)) + 
#   labs(title = "R-Squared for Occupation (Most Detailed)\nRegressed on Log(Income), all Cases") + 
#   facet_wrap(~sex)

```

### Drop missing + 0 earners

```{r}

acs2 <- acs2[
  # acs2$empstat == "at work" &
  #                            (acs2$schlcoll %like% "does not attend|niu"| is.na(acs2$schlcoll)) &
  #                            as.numeric(as.character(acs2$age)) %in% 25:64 &
                             #acs2$wkswork1 >= 30 & ### !!!!! ahhhh
                             acs2$incwage != 99999999 &
                             acs2$incwage > 0,]

r_sq_dt <- data.table()
i <- 0
for(c.year in unique(acs2$year)){
  for(c.educ_yr in unique(acs2[!is.na(ed_categ)]$ed_categ)){
    for(c.sex in list("male", "female", c("male", "female"))){
      i <- i + 1
      print(i)
      out <- lm(log_incwage ~ as.factor(occ2010), data = acs2[year == c.year & ed_categ == c.educ_yr &
                                                       sex %in% c.sex])
      out_dt <- data.table(year = c.year, ed_categ = c.educ_yr,
                           sex = paste0(c.sex, collapse = ","),
                           r_sq = summary(out)$r.squared)
      r_sq_dt <- rbind(r_sq_dt, out_dt, fill = T)
    }
  }
}

ggplot(r_sq_dt) +
  geom_point(aes(x = year, y = r_sq, color = ed_categ)) +
  geom_line(aes(x = year, y = r_sq, color = ed_categ, group = ed_categ)) +
  labs(title = "R-Squared for Occupation (Most Detailed)\nRegressed on Log(Income), all Cases") +
  facet_wrap(~sex)

```

### Drop missing + 0 earners and part timers

```{r}

acs2 <- acs2[
  # acs2$empstat == "at work" &
  #                            (acs2$schlcoll %like% "does not attend|niu"| is.na(acs2$schlcoll)) &
  #                            as.numeric(as.character(acs2$age)) %in% 25:64 &
                             acs2$wkswork1 >= 30 & ### !!!!! ahhhh
                             acs2$incwage != 99999999 &
                             acs2$incwage > 0,]

r_sq_dt <- data.table()
i <- 0
for(c.year in unique(acs2$year)){
  for(c.educ_yr in unique(acs2[!is.na(ed_categ)]$ed_categ)){
    for(c.sex in list("male", "female", c("male", "female"))){
      i <- i + 1
      print(i)
      out <- lm(log_incwage ~ as.factor(occ2010), data = acs2[year == c.year & ed_categ == c.educ_yr & 
                                                       sex %in% c.sex])
      out_dt <- data.table(year = c.year, ed_categ = c.educ_yr,
                           sex = paste0(c.sex, collapse = ","),
                           r_sq = summary(out)$r.squared)
      r_sq_dt <- rbind(r_sq_dt, out_dt, fill = T)
    }
  }
}

ggplot(r_sq_dt) + 
  geom_point(aes(x = year, y = r_sq, color = ed_categ)) + 
  geom_line(aes(x = year, y = r_sq, color = ed_categ, group = ed_categ)) + 
  labs(title = "R-Squared for Occupation (Most Detailed)\nRegressed on Log(Income), all Cases") + 
  facet_wrap(~sex)

```

### Compositional Biases induced by dropping part time workers and 0 earners + missing

```{r}
acs2 <- acs2[
   acs2$empstat %like% "at work|armed|has job" &
                              (acs2$schlcoll %like% "does not attend|niu"| is.na(acs2$schlcoll)) &
                              as.numeric(as.character(acs2$age)) %in% 25:64 &
                             acs2$wkswork1 >= 30 & ### !!!!! ahhhh
                             acs2$incwage != 99999999 &
                             acs2$incwage > 0,]


r_sq_dt <- data.table()
i <- 0
for(c.year in unique(acs2$year)){
  for(c.educ_yr in unique(acs2[!is.na(ed_categ)]$ed_categ)){
    for(c.sex in list("male", "female", c("male", "female"))){
      i <- i + 1
      print(i)
      out <- lm(log_incwage ~ as.factor(occ2010), data = acs2[year == c.year & ed_categ == c.educ_yr & 
                                                       sex %in% c.sex])
      out_dt <- data.table(year = c.year, ed_categ = c.educ_yr,
                           sex = paste0(c.sex, collapse = ","),
                           r_sq = summary(out)$r.squared)
      r_sq_dt <- rbind(r_sq_dt, out_dt, fill = T)
    }
  }
}

ggplot(r_sq_dt) + 
  geom_point(aes(x = year, y = r_sq, color = ed_categ)) + 
  geom_line(aes(x = year, y = r_sq, color = ed_categ, group = ed_categ)) + 
  labs(title = "R-Squared for Occupation (Most Detailed)\nRegressed on Log(Income), Partial Cases") + 
  facet_wrap(~sex)

```
