---
title: "analysis_III_time_series_abs"
author: "Hunter York"
date: "10/26/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(ggplot2)
library(dplyr)
library(magrittr)
library(stringr)
library(readxl)
library(ggrepel)
library(foreign)
library(ipumsr)
library(data.table)



options
theme_set(theme_bw())

# load the data
setwd("/Users/hyork/Documents/projects/occupation/code")

weighted.var <- function(x, w, na.rm = FALSE) {
    if (na.rm) {
        w <- w[i <- !is.na(x)]
        x <- x[i]
    }
    sum.w <- sum(w)
    sum.w2 <- sum(w^2)
    mean.w <- sum(x * w) / sum(w)
    (sum.w / (sum.w^2 - sum.w2)) * sum(w * (x - mean.w)^2, na.rm =
na.rm)
}


# #census_1940 <- readRDS("../inputs/usa_00005.rds")
# census_1940 <- readstata13::read.dta13("../inputs/usa_00009.dta")
# census_1940 <- data.table(census_1940)
# #keep vars
# census_1940 <- census_1940[,.(race, hispan, year,statefip, countyicp,metro,perwt, sex, age, school,educ, empstat,occ,occ1950, ind,ind1950, classwkr, wkswork2, incwage)]
# 
# # subset
# census_1940 <- census_1940[empstat == "employed" &
#                              (school == "no, not in school"| as.numeric(as.character(age)) >=34) &
#                              as.numeric(as.character(age)) %in% 25:64 &
#                             wkswork2 %like% "40|48|50" &
#                              incwage != 999998 &
#                              incwage > 0]

# #saveRDS(census_1940, "../inputs/usa_00009.rds")
census_1940 <- readRDS( "../inputs/usa_00009.rds")

setnames(census_1940, c("occ1950", "ind1950"), c("origocc1950", "origind1950"))
# derive a 5-year age var
census_1940[, age_cat := paste0(floor(as.numeric(as.character(age))/5)*5,
                                " to ", 
                                floor(as.numeric(as.character(age))/5)*5 + 4)]

# make occ var char
census_1940[, occ := as.character(occ)]

# 
census_1940[nchar(occ) == 1, occ := paste0("00", occ)]
census_1940[nchar(occ) == 2, occ := paste0("0", occ)]

# make first dig var
census_1940[, first_dig := substr(occ, 1,1)]

# make map for first dig categories
map <- data.table(first_dig = as.character(0:9), 
                  occ_categ = c("Professional", "Clerical", "Service",
                                "Agricultural, etc.", "Skilled", "Skilled",
                                "Semiskilled", "Semiskilled", "Unskilled",
                                "Unskilled"))
census_1940 <- merge(census_1940, map, by = "first_dig")

# merge census 1940 file on ind codes
ind_codes <- read_xlsx("../ref/IND1940.xlsx") %>% data.table()
ind_codes[, IND := as.numeric(IND)]
census_1940 <- merge(census_1940, 
                     ind_codes[!is.na(IND) & IND != 996], 
                     by.x = "ind", by.y = "IND", all.x = T)

# crosswalk to ppp
ppp <- read_excel("../ref/CPI_U_RS.xlsx") %>% data.table
ppp[1, year := 1940]
ppp[1, cpi := 32]
ppp[, year := as.character(year)]
ppp[, xwalk_fac := cpi/369.8]
census_1940 <- merge(census_1940, ppp, by = "year")
census_1940[, incwage := incwage/xwalk_fac]
# create log inc wage
census_1940[, log_incwage := log(incwage + 1)]
```

# Crosswalk

```{r}
# load in xwalk
xwalk <- data.table(read_excel("../ref/Census_integrated_occ_crosswalks.xlsx"))
xwalk_long <- melt(xwalk, id.vars = c("OCC1950", "Occupation category description"))
setnames(xwalk_long, c("OCC1950", "OCC1950_desc", "year", "orig_occ"))
xwalk_long[as.character(year) =="ACS 2000-02", year := "2000ACS"]
xwalk_long[as.character(year)  == "ACS 2003-", year := "2018"]

# copy 1950 vals to 1940 for nowxwalk_long
xwalk_long[, year := as.character(year)]
xwalk_long[year == 1950] %>% 
  .[, year := 1940] %>% 
  rbind(., xwalk_long) -> temp
xwalk_long[, orig_occ := as.numeric(orig_occ)]
census_1940[, occ := as.numeric(occ)]
# merge on census
census_1940 <- merge(census_1940, xwalk_long, by.y = c("year", "orig_occ"), by.x = c("year", "occ"), all.x = T)

get_vars <- function(c.data, c.by_vars,c.by_vars_2, c.var_interest){
  out_dt_1 <- c.data[,.(w_i_ss = weighted.var(get(c.var_interest), perwt) * .N,
                        N = .N,
                        k = max(.GRP)),
                     by = c(c.by_vars, c.by_vars_2)]
  out_2 <- c.data[,.(tot_ss=weighted.var(get(c.var_interest), perwt) * .N), by = c.by_vars_2]
  out_dt_1 <- merge(out_dt_1, out_2, by = c.by_vars_2)
  out_dt_1 <- out_dt_1[!is.na(tot_ss)& !is.na(w_i_ss) & !is.nan(tot_ss)& !is.nan(w_i_ss) & 
                          !is.infinite(tot_ss)& !is.infinite(w_i_ss),
                       .(avg_within_var = sum(w_i_ss),
                         avg_total_var = mean(tot_ss),
                         avg_between_var = mean(tot_ss) -sum(w_i_ss), 
                         N = sum(N),
                         k = length(unique(N[!is.na(w_i_ss)]))),
                       by = c.by_vars_2]
  return(out_dt_1)
}

dem_var_gettr2 <- function(c.dat2, c.by_vars_2){
  occ_only <- get_vars(c.dat2,
                       c.by_vars = c("origocc1950"),
                       c.by_vars_2 = c.by_vars_2,
                       c.var_interest = "log_incwage")
  occ_only[, grouping := "Occupation"]
  ind_only <- get_vars(c.dat2,
                       c.by_vars = c("origind1950"),
                       c.by_vars_2 = c.by_vars_2,

                       c.var_interest = "log_incwage")
    ind_only[, grouping := "Industry"]

  occ_ind <- get_vars(c.dat2,
                      c.by_vars = c("origocc1950", "origind1950"),
                      c.by_vars_2 = c.by_vars_2,
                      c.var_interest = "log_incwage")
    occ_ind[, grouping := "Occ + Ind"]

  
  out_dt <- rbindlist(list(occ_only, ind_only, occ_ind ))
  return(out_dt)
}


temp <- dem_var_gettr2(census_1940, c.by_vars_2 = c("year"))


plot_dt2 <- temp

plot_dt2[, f_stat := (avg_between_var/(k-1))/((avg_within_var)/(N-k))]
plot_dt2[, ms_bw := (avg_between_var/(k-1))]
plot_dt2[, ms_wi := (avg_within_var/(N-k))]

plot_dt2[, within_perc :=
          avg_within_var/
          (avg_within_var+avg_between_var)]
plot_dt2[, between_perc :=
          avg_between_var/
          (avg_within_var+avg_between_var)]
plot_dt2[, bw_wi_perc_ratio := between_perc/within_perc]
plot_dt2[, total_var := avg_total_var/N]
plot_dt2[, within_var := avg_within_var/N]
plot_dt2[, between_var := avg_between_var/N]


# cast long
plot_dt2_long <- melt(plot_dt2, id.vars = c("year",
                                          "grouping"),
                     measure.vars = c("within_perc",
                                      "between_perc",
                                      "between_var",
                                      "bw_wi_perc_ratio",
                                      "total_var",
                                      "within_var",
                                      "ms_wi",
                                      "ms_bw"))

gg2 <- ggplot(plot_dt2_long[variable %like% "between_perc|total_var|within_var|between_var"]) +
  geom_line(aes(x = year, y = value,
                color = grouping, group = grouping))+
  facet_wrap(~variable)+
  labs(x = "Year", y = "Percent Total Variance Explained\nBy Between-Group Variation",
       title = "Results using xwalked OCC and 1950IND"
       )
print(gg2)
```

# Task 2 - Use some specific occupations to examine speciic trajectories

I'm using six relatively stable single occupational designations from the 1950 codings: teachers, industrial engineers, waiters/waitresses, accountants/auditors, lawyers and judges, clergymen.

```{r}
census_subset <- census_1940[OCC1950 %in% c(0, 784, 55, 45, 9, 93) & !is.na(origocc1950)]

# calculate variance by job, year
vars_by_job_year <- census_subset[,.(variance = var(log_incwage),
                                   median_wage = median(log_incwage),
                                   gini = DescTools::Gini(log_incwage), 
                                   count = .N), by = .(OCC1950, year)]

# merge on descs
temp <- data.table(OCC1950 =as.character(c(0, 784, 55, 45, 9, 93)), OCC1950_desc_std = c("Accountants", "Waiters", "Lawyers", "Engineers", "Clergy", "Teachers(NEC)"))

vars_by_job_year <- merge(vars_by_job_year, temp)

# plot
vars_by_job_year %>% melt(., id.vars = c("OCC1950_desc_std", "OCC1950", "year")) %>% 
ggplot(.) + 
  geom_line(aes(x = year, y = value, group = OCC1950, color = OCC1950_desc_std)) + 
  facet_wrap(~variable, scales = "free_y")
```

# Task 3 - do the above analysis for all occupations and see if there are any major outliers in terms of increased or decrease equality

## Increased equality

```{r}

# calculate variance by job, year
vars_by_job_year <- census_1940[,.(variance = var(log_incwage),
                                   median_wage = median(log_incwage),
                                   gini = DescTools::Gini(log_incwage), 
                                   count = .N), by = .(OCC1950, year)]

vars_by_job_year[,gini_diff := max(gini) - min(gini), by = .(OCC1950)]
vars_by_job_year[, median_count := median(count[year <= 2000]), by = .(OCC1950)]

# merge on descs
temp <- data.table(OCC1950 =as.character(c(57, 97, 100, 595, 764, 781, 820, 999)), OCC1950_desc_std = c("Musicians/Music Teachers", "Therapists & Healers(NEC)", "Farmers", "Members of Armed Services", "Housekeepers, except private", "Practical Nurses", "Farm laborers", "NA"))

vars_by_job_year <- merge(vars_by_job_year, temp)

# plot
vars_by_job_year %>% melt(., id.vars = c("OCC1950_desc_std", "OCC1950", "year")) %>% 
  .[variable %like%"var|wage|^gini$"] %>% 
ggplot(.) + 
  geom_line(aes(x = year, y = value, group = OCC1950, color = OCC1950_desc_std)) + 
  facet_wrap(~variable, scales = "free_y")

```

## Decreased equality

```{r}
# calculate variance by job, year
vars_by_job_year <- census_1940[,.(variance = var(log_incwage),
                                   median_wage = median(log_incwage),
                                   gini = DescTools::Gini(log_incwage), 
                                   count = .N), by = .(OCC1950, year)]

vars_by_job_year[,gini_diff := gini[year == min(year)] - gini[year == max(year)], by = .(OCC1950)]
vars_by_job_year[, median_count := median(count[year <= 2000]), by = .(OCC1950)]
vars_by_job_year <- vars_by_job_year[gini_diff < -0 & median_count > 500]

# merge on descs
temp <- data.table(OCC1950 =as.character(c(45, 750, 0, 81, 551, 682)), OCC1950_desc_std = c("Industrial Engineers", "Bartenders", "Accountants and Auditors", "Office Machine Mechanics", "Taxi Drivers/Chauffers"))

vars_by_job_year <- merge(vars_by_job_year, temp)

# plot
vars_by_job_year %>% melt(., id.vars = c("OCC1950_desc_std", "OCC1950", "year")) %>% 
  .[variable %like%"var|wage|^gini$"] %>% 
ggplot(.) + 
  geom_line(aes(x = year, y = value, group = OCC1950, color = OCC1950_desc_std)) + 
  facet_wrap(~variable, scales = "free_y")
```

## Aggregate up to less granular job titles using census hierarchy and repeat

```{r}
# create crosswalk for this
base_xwalk <- read_excel("../ref/Census_integrated_occ_crosswalks.xlsx") %>% data.table()
base_xwalk <- base_xwalk[,.(OCC1950, `Occupation category description`)]
base_xwalk[OCC1950 == "#",parent_occ1950 := `Occupation category description`]
for(i in 2:nrow(base_xwalk)){
  if (is.na(base_xwalk[i, parent_occ1950])){
    base_xwalk[i, parent_occ1950 := base_xwalk[i-1, parent_occ1950]]
  }
}
census_1940 <- merge(census_1940, unique(base_xwalk[OCC1950 != "#",.(OCC1950, parent_occ1950)]), by = "OCC1950", all.x = T)


# calculate variance by job, year
vars_by_job_year <- census_1940[,.(variance = var(log_incwage),
                                   median_wage = median(log_incwage),
                                   gini = DescTools::Gini(log_incwage), 
                                   count = .N), by = .(parent_occ1950, year)]



# plot
vars_by_job_year %>% melt(., id.vars = c("parent_occ1950", "year")) %>% 
ggplot(.) + 
  geom_line(aes(x = year, y = value, group = parent_occ1950, color = parent_occ1950)) + 
  facet_wrap(~variable, scales = "free_y")

```

# Characterize within and between occupation heterogeneity within each of these groupings

```{r}
temp <- dem_var_gettr2(census_1940, c.by_vars_2 = c("year", "parent_occ1950"))


plot_dt2 <- temp

plot_dt2[, f_stat := (avg_between_var/(k-1))/((avg_within_var)/(N-k))]
plot_dt2[, ms_bw := (avg_between_var/(k-1))]
plot_dt2[, ms_wi := (avg_within_var/(N-k))]

plot_dt2[, within_perc :=
          avg_within_var/
          (avg_within_var+avg_between_var)]
plot_dt2[, between_perc :=
          avg_between_var/
          (avg_within_var+avg_between_var)]
plot_dt2[, bw_wi_perc_ratio := between_perc/within_perc]
plot_dt2[, total_var := avg_total_var/N]
plot_dt2[, within_var := avg_within_var/N]
plot_dt2[, between_var := avg_between_var/N]


# cast long
plot_dt2_long <- melt(plot_dt2, id.vars = c("year",
                                          "grouping",
                                          "parent_occ1950"),
                     measure.vars = c("within_perc",
                                      "between_perc",
                                      "between_var",
                                      "bw_wi_perc_ratio",
                                      "total_var",
                                      "within_var",
                                      "ms_wi",
                                      "ms_bw"))

gg2 <- ggplot(plot_dt2_long[variable %like% "between_perc|total_var|within_var|between_var" & grouping == "Occupation" & parent_occ1950 != "Farmers:"]) +
  geom_line(aes(x = year, y = value,
                color = parent_occ1950, group = parent_occ1950))+
  facet_wrap(~variable, scales = "free")+
  labs(x = "Year", y = "Value",
       title = "Heterogeneity Explained Within Census Category by Occupation"
       )
```

## Repeat by education

```{r}
census_1940[as.numeric(educ) %in% 0:7, ed := "Less than High School"]
census_1940[as.numeric(educ)  %in% 8:10, ed := "Some College"]
census_1940[as.numeric(educ)  %in% 11, ed := "4-Year Degree"]
census_1940[as.numeric(educ)  %in% 12, ed := "Post-Bachelors"]
census_1940[, ed := factor(ed, levels = c("Less than High School","Some College","4-Year Degree", "Post-Bachelors")
)]


temp <- dem_var_gettr2(census_1940, c.by_vars_2 = c("year", "parent_occ1950", "ed"))


plot_dt2 <- temp

plot_dt2[, f_stat := (avg_between_var/(k-1))/((avg_within_var)/(N-k))]
plot_dt2[, ms_bw := (avg_between_var/(k-1))]
plot_dt2[, ms_wi := (avg_within_var/(N-k))]

plot_dt2[, within_perc :=
          avg_within_var/
          (avg_within_var+avg_between_var)]
plot_dt2[, between_perc :=
          avg_between_var/
          (avg_within_var+avg_between_var)]
plot_dt2[, bw_wi_perc_ratio := between_perc/within_perc]
plot_dt2[, total_var := avg_total_var/N]
plot_dt2[, within_var := avg_within_var/N]
plot_dt2[, between_var := avg_between_var/N]


# cast long
plot_dt3_long <- melt(plot_dt2, id.vars = c("year",
                                          "grouping",
                                          "parent_occ1950",
                                          "ed",
                                          "N"),
                     measure.vars = c("within_perc",
                                      "between_perc",
                                      "between_var",
                                      "bw_wi_perc_ratio",
                                      "total_var",
                                      "within_var",
                                      "ms_wi",
                                      "ms_bw"))

plot_dt3_long <- plot_dt3_long[,.(value = weighted.mean(value, N)), by = .(year, grouping, parent_occ1950, variable)]

#merge on dataset that didn't control for ed
setnames(plot_dt2_long, "value", "value_uncontrolled")
plot_dt3_long <- merge(plot_dt3_long, plot_dt2_long)
plot_dt3_long[, value_ratio := value/value_uncontrolled]


gg3 <- ggplot(plot_dt3_long[variable %like% "between_perc|total_var|within_var|between_var" & grouping == "Occupation" & parent_occ1950 != "Farmers:"]) +
  geom_line(aes(x = year, y = value,
                color = parent_occ1950, group = parent_occ1950))+
  facet_wrap(~variable, scales = "free")+
  labs(x = "Year", y = "Value",
       title = ""
       )
print(gg3)
gg4 <- ggplot(plot_dt3_long[variable %like% "between_perc" & grouping == "Occupation" & parent_occ1950 != "Farmers:"]) + 
    geom_line(aes(x = year, y = value_ratio,
                color = parent_occ1950, group = parent_occ1950))+
  facet_wrap(~variable, scales = "free")+
  labs(x = "Year", y = "Value",
       title = "Attenuation of Explained Variance When Controlling for Education"
       )
print(gg4)
```

# See which occupations adding industry helps explain variation

```{r}

temp <- dem_var_gettr2(census_1940, c.by_vars_2 = c("year", "OCC1950"))


plot_dt2 <- temp

plot_dt2[, f_stat := (avg_between_var/(k-1))/((avg_within_var)/(N-k))]
plot_dt2[, ms_bw := (avg_between_var/(k-1))]
plot_dt2[, ms_wi := (avg_within_var/(N-k))]

plot_dt2[, within_perc :=
          avg_within_var/
          (avg_within_var+avg_between_var)]
plot_dt2[, between_perc :=
          avg_between_var/
          (avg_within_var+avg_between_var)]
plot_dt2[, bw_wi_perc_ratio := between_perc/within_perc]
plot_dt2[, total_var := avg_total_var/N]
plot_dt2[, within_var := avg_within_var/N]
plot_dt2[, between_var := avg_between_var/N]


# cast long
plot_dt2_long <- melt(plot_dt2, id.vars = c("year",
                                          "grouping",
                                          "OCC1950"),
                     measure.vars = c("within_perc",
                                      "between_perc",
                                      "between_var",
                                      "bw_wi_perc_ratio",
                                      "total_var",
                                      "within_var",
                                      "ms_wi",
                                      "ms_bw"))
plot_dt2_long[variable %like% "between_perc" & grouping == "Industry" & value > .5, unique(OCC1950)] -> temp
gg2 <- ggplot(plot_dt2_long[variable %like% "between_perc" & grouping == "Industry" & OCC1950 %in% temp]) +
  geom_line(aes(x = year, y = value,
                color = OCC1950, group = OCC1950))+
  facet_wrap(~variable)+
  labs(x = "Year", y = "Percent Total Variance Explained\nBy Between-Group Variation",
       title = "Results using xwalked OCC and 1950IND"
       )

print(gg2)

census_1940[OCC1950 %in% temp, .(OCC1950, OCC1950_desc)] %>% unique()
```