---
title: "analysis_i_variance_decomp"
author: "Hunter York"
date: "9/27/2020"
output: pdf_document
---

---
title: "analysis_i_variance_decomp"
author: "Hunter York"
date: "9/27/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DescTools)
library(data.table)
library(ggplot2)
library(dplyr)
library(magrittr)
library(stringr)
library(readxl)
library(ggrepel)
library(foreign)
library(ipumsr)


options
theme_set(theme_bw())

# load the data
setwd("/Users/hyork/Documents/projects/occupation/code")



# #census_1940 <- readRDS("../inputs/usa_00005.rds")
# census_1940 <- readstata13::read.dta13("../inputs/usa_00006.dta")
# census_1940 <- data.table(census_1940)
# # keep vars
# census_1940 <- census_1940[,.(statefip, countyicp, urban, farm, sex, age, race, hispan, school, higrade, empstat,occ, ind, classwkr, wkswork1, hrswork1, incwage)]
# 
# # subset
# census_1940 <- census_1940[empstat == "employed" &
#                              school == "no, not in school" &
#                              age %in% 25:64 &
#                              hrswork1 >= 30 &
#                              wkswork1 >= 40 &
#                              incwage != 999998 &
#                              incwage > 0]

#saveRDS(census_1940, "../inputs/usa_00006.rds")
census_1940 <- readRDS( "../inputs/usa_00006.rds")

# derive a income/hr worked per week var
census_1940[, income_per_hr := incwage/hrswork1]

# derive a 5-year age var
census_1940[, age_cat := paste0(floor(as.numeric(as.character(age))/5)*5,
                                " to ", 
                                floor(as.numeric(as.character(age))/5)*5 + 4)]

# make occ var char
census_1940[, occ := as.character(occ)]

# 
census_1940[nchar(occ) == 1, occ := paste0("00", occ)]
census_1940[nchar(occ) == 2, occ := paste0("0", occ)]

# make first dig var
census_1940[, first_dig := substr(occ, 1,1)]

# make map for first dig categories
map <- data.table(first_dig = as.character(0:9), 
                  occ_categ = c("Professional", "Clerical", "Service",
                                "Agricultural, etc.", "Skilled", "Skilled",
                                "Semiskilled", "Semiskilled", "Unskilled",
                                "Unskilled"))
census_1940 <- merge(census_1940, map, by = "first_dig")

# merge census 1940 file on ind codes
ind_codes <- read_xlsx("../ref/IND1940.xlsx") %>% data.table()
ind_codes[, IND := as.numeric(IND)]
census_1940 <- merge(census_1940, 
                     ind_codes[!is.na(IND) & IND != 996], 
                     by.x = "ind", by.y = "IND", all.x = T)

# create log inc wage
census_1940[, log_incwage := log(incwage + 1)]
```

# Step 1 - Decompose variance by industry, by occupation, and then by their intersection

## Step 1.2 - Quick descriptive table of Occupation + Industry intersection

There are 24942 intersections of occupation and industry. (230 occupations * 134 industry) - around 5000 missing combinations. About 5000 combos only have one respondent in the category. 4725 have 50 or more respondents, and 1397 have more than 1000. 83% of respondents are captured in all occ+ind combos with at least 5000 respondants, and 94% in all combos with at least 1000 respondents. 

```{r}
census_1940 %>%
  .[,.N, by = .(occ, ind)] %>% 
  .[order(N)] %>% 
  .[!is.na(N)] %>% 
   ggplot(.) + 
     geom_density(aes(x= N)) +
  scale_x_continuous(trans = "log10") + 
  labs(x = "Number of Individuals in each Occ + Ind combo",
       title = "Distribution of Occupation and Industry Intersections") 
```

## Step 2 - Between occupation vs within occupation variance, and then for industry and their intersection

This following section uses the following formula to calculate the statistics used in this section.

$$SS_{Total} = \sum_{j=1}^k{\sum_{i=1}^{n_j}{(x_{ij}-\bar{x}^2)}}$$
$$SS_{Between} = \sum{n_j(\bar{x_j} - \bar{x})^2}$$

$$SS_{Within} = SS_{Total} - SS_{Between}$$

$$MS_{Between} = \frac{SS_{Between}}{k-1}$$
$$MS_{Within} = \frac{SS_{Within}}{n-k}$$

$$F = \frac{MS_{Between}}{MS_{Within}}$$
## Replication of Xie & Killewald

$$\eta^2 = \frac{SS_{Between}}{SS_{Total}}$$
**NB: This is equivalent to the method used by Xie & Killewald wherein $\eta^2$ is what would be returned if a linear model were fit and ANOVA was run on it.**





```{r}
# create function to get var from a variable
get_vars <- function(c.data, c.by_vars, c.var_interest){
  out_dt_1 <- c.data[,.(w_i_ss = var(get(c.var_interest)) * .N,
                        N = .N,
                        k = max(.GRP)),
                     by = c.by_vars]
  out_dt_1[,tot_ss := sum((c.data[,(get(c.var_interest))])^2) - 
             (((sum(c.data[,(get(c.var_interest))]))^2)/
                length(c.data[,(get(c.var_interest))]))]
  out_dt_1 <- out_dt_1[!is.na(tot_ss)& !is.na(w_i_ss),
                       .(avg_within_var = sum(w_i_ss),
                         avg_total_var = mean(tot_ss),
                         avg_between_var = mean(tot_ss) -sum(w_i_ss), 
                         N = sum(N),
                         k = length(unique(N[!is.na(w_i_ss)])))]
  return(out_dt_1)
}

#  create another function to loop over data and 
#  calculate occ, ind, and occ + ind var

dem_var_gettr <- function(c.dat2,
                          c.age_cat,
                          c.sex,
                          c.urban){
  occ_only <- get_vars(c.dat2[age_cat %in% c.age_cat &
                                sex %in% c.sex &
                                urban %in% c.urban],
                       c.by_vars = "occ",
                       c.var_interest = "log_incwage")
  ind_only <- get_vars(c.dat2[age_cat %in% c.age_cat &
                                sex %in% c.sex&
                                urban %in% c.urban],
                       c.by_vars = "ind",
                       c.var_interest = "log_incwage")
  occ_ind <- get_vars(c.dat2[age_cat %in% c.age_cat &
                               sex %in% c.sex&
                               urban %in% c.urban],
                      c.by_vars = c("occ", "ind"), 
                      c.var_interest = "log_incwage")
  
  out_dt <- rbindlist(list(occ_only, ind_only, occ_ind ))
  out_dt[,grouping := c("Occupation", "Industry", "Ind. + Occ.")]
  out_dt[, age_cat := paste(c.age_cat, collapse = ", ")]
  out_dt[, sex :=  paste(c.sex, collapse = ", ")]
  out_dt[, urban :=  paste(c.urban, collapse = ", ")]
  
  return(out_dt)
}

temp_male_urban <- lapply(unique(census_1940$age_cat), dem_var_gettr, 
                          c.dat = census_1940,
                          c.sex = "male",
                          c.urban = "urban")
temp_male_rural <- lapply(unique(census_1940$age_cat), dem_var_gettr, 
                          c.dat = census_1940,
                          c.sex = "male",
                          c.urban = "rural")

temp_female_urban <- lapply(unique(census_1940$age_cat), dem_var_gettr, 
                            c.dat = census_1940,
                            c.sex = "female",
                            c.urban = "urban")
temp_female_rural <- lapply(unique(census_1940$age_cat), dem_var_gettr, 
                            c.dat = census_1940,
                            c.sex = "female",
                            c.urban = "rural")

temp_female_both <- lapply(unique(census_1940$age_cat), dem_var_gettr, 
                            c.dat = census_1940,
                            c.sex = "female",
                            c.urban = c("urban", "rural"))
temp_male_both <- lapply(unique(census_1940$age_cat), dem_var_gettr, 
                            c.dat = census_1940,
                            c.sex = "male",
                            c.urban = c("urban", "rural"))

plot_dt <- rbindlist(list(rbindlist(temp_male_rural), 
                          rbindlist(temp_female_rural),
                          rbindlist(temp_male_urban),
                          rbindlist(temp_female_urban),
                          rbindlist(temp_male_both),
                          rbindlist(temp_female_both)))

plot_dt[, f_stat := (avg_between_var/(k-1))/((avg_within_var)/(N-k))]
plot_dt[, ms_bw := (avg_between_var/(k-1))]
plot_dt[, ms_wi := (avg_within_var/(N-k))]

plot_dt[, age_start := as.numeric(substr(age_cat,1,2))]

plot_dt[, within_perc :=
          avg_within_var/
          (avg_within_var+avg_between_var)]
plot_dt[, between_perc :=
          avg_between_var/
          (avg_within_var+avg_between_var)]
plot_dt[, bw_wi_perc_ratio := between_perc/within_perc]

# ggplot(plot_dt)+ 
#   geom_line(aes(x = age_start, y = f_stat, color = grouping)) +
#   facet_grid(urban~sex) +
#   geom_hline(yintercept = 1, linetype = "dashed")

# cast long
plot_dt_long <- melt(plot_dt, id.vars = c("age_cat",
                                          "urban",
                                          "grouping",
                                          "sex",
                                          "age_start"),
                     measure.vars = c("within_perc",
                                      "between_perc",
                                      "bw_wi_perc_ratio",
                                      "ms_wi",
                                      "ms_bw"))

ggplot(plot_dt_long[variable %like% "between_perc"]) +
  geom_line(aes(x = age_start, y = value,
                color = grouping))+
  facet_grid(urban ~ sex) + 
  labs(x = "Age", y = "Percent Total Variance Explained")
# 
# ggplot(plot_dt_long[variable %in% c("ms_wi", "ms_bw")]) + 
#   geom_line(aes(x = age_start, y = value,
#                 color = grouping,
#                 linetype = variable))+ 
#   facet_grid(urban ~ sex)


```

# Step 2 - Try to recreate Xie, Killewald, and Near 2016

## Step 1 - Bin by Education and repeat above analysis

```{r}
# create education category
census_1940[higrade %in% 0:11, ed := "Less than High School"]
census_1940[higrade %in% 12:15, ed := "Some College"]
census_1940[higrade %in% 16, ed := "4-Year Degree"]
census_1940[higrade %in% 17:80, ed := "Post-Bachelors"]
census_1940[, ed := factor(ed, levels = c("Less than High School","Some College","4-Year Degree", "Post-Bachelors")
)]
#  create another function to loop over data and 
#  calculate occ, ind, and occ + ind var

dem_var_gettr_ed <- function(c.dat2,
                             c.age_cat,
                             c.sex,
                             c.ed){
  occ_only <- get_vars(c.dat2[age_cat %in% c.age_cat &
                                sex %in% c.sex &
                                ed %in% c.ed],
                       c.by_vars = "occ",
                       c.var_interest = "log_incwage")
  ind_only <- get_vars(c.dat2[age_cat %in% c.age_cat &
                                sex %in% c.sex&
                                 ed %in% c.ed],
                       c.by_vars = "ind",
                       c.var_interest = "log_incwage")
  occ_ind <- get_vars(c.dat2[age_cat %in% c.age_cat &
                               sex %in% c.sex&
                                ed %in% c.ed],
                      c.by_vars = c("occ", "ind"), 
                      c.var_interest = "log_incwage")
  
  out_dt <- rbindlist(list(occ_only, ind_only, occ_ind ))
  out_dt[,grouping := c("Occupation", "Industry", "Ind. + Occ.")]
  out_dt[, age_cat := paste(c.age_cat, collapse = ", ")]
  out_dt[, sex :=  paste(c.sex, collapse = ", ")]
  out_dt[, ed :=  paste(c.ed, collapse = ", ")]
  
  return(out_dt)
}

ed_gettr <- function(c.ed2){
  temp <- lapply(unique(census_1940$age_cat), dem_var_gettr_ed, 
                 c.dat = census_1940,
                 c.sex = "male",
                 c.ed = c.ed2)
  temp2 <- lapply(unique(census_1940$age_cat), dem_var_gettr_ed, 
                 c.dat = census_1940,
                 c.sex = "female",
                 c.ed = c.ed2)
  temp <- c(temp, temp2)
  return(temp)
}


lapply(
  unique(census_1940[!is.na(ed)]$ed),
                            ed_gettr) %>% 
  lapply(., rbindlist) %>% 
  rbindlist() -> plot_dt2

plot_dt2[, f_stat := (avg_between_var/(k-1))/((avg_within_var)/(N-k))]
plot_dt2[, age_start := as.numeric(substr(age_cat,1,2))]

plot_dt2[, ed := factor(ed, levels = c("Less than High School","Some College","4-Year Degree", "Post-Bachelors")
)]

plot_dt2[, r_squared := 1 - avg_within_var/(avg_within_var + avg_between_var)]

plot_dt2[, between_perc :=
          avg_between_var/
          (avg_within_var+avg_between_var)]

ggplot(plot_dt2)+ 
  geom_line(aes(x = age_start, y = between_perc, color = grouping)) +
  facet_grid(ed~sex) +
    labs(x = "Age", y = "Percent Total Variance Explained")


ggplot(plot_dt2[age_start %in% c(25, 35,45,55)])+ 
  geom_line(aes(x = ed, y = between_perc, color = grouping, group = grouping)) +
  facet_grid(age_cat~sex) +
  theme(axis.text.x =  element_text(angle = 90))+
    labs(x = "Education", y = "Percent Total Variance Explained")



```


